---
title: Scottish Contact Survey Demographics Summary
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

source(here::here("code", "00_setup.R"))

library(ggplot2)
library(scsplot)
library(flextable)
library(officer)

knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      fig.width = 15, fig.width = 10)

theme_set(scs_theme())
dodge <- position_dodge(width = 0.9)

```

```{r opt-outs-function, include = FALSE}

opt_out_data <- function(x) {

  get_opt_outs <- function(y) {
    
    file <- suppressWarnings(
      scs_filepath("opt-outs-anon", y, create_folder = FALSE)
    )
    
    if(!file.exists(file)) {
      stop("File does not exist: ", file)
    }
    
    read_rds(file) %>%
      filter(gender %in% c("Male", "Female")) %>%
      group_by(age_group, gender) %>%
      summarise(n = n(), .groups = "drop")
    
  }
  
  if(!(x %in% 30:42) & !(x >= 44)) {
    stop("Wave must be greater or equal to 30, and not 43.")
  }

  data <- if(x %in% 30:42) {
    bind_rows(get_opt_outs(paste0(x, "A")), get_opt_outs(paste0(x, "B")))
  } else {
    get_opt_outs(x)
  }

  data %>%
    group_by(age_group, gender) %>%
    summarise(across(everything(), sum), .groups = "drop") %>%
    mutate(
      survey_wave = x,
      date = start_date(as.numeric(x), "A")
    ) %>%
    select(survey_wave, date, everything())

}

```

<br>

## Wave `r wave`

Wave `r wave` of the Scottish Contact Survey was open from `r date_range(wave, year = TRUE)`.

```{r data-latest, include = FALSE}

# Age and gender of responses to latest wave
resp <- 
  scs_filepath("registration-data-anon", wave) %>%
  read_rds() %>%
  mutate(age_group = age(date_of_birth, wave, grouped = TRUE)) %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(age_group, gender) %>%
  summarise(n_resp = n(), .groups = "drop")

# Age and gender of those invited to latest wave
reg <-
  scs_filepath("registration-data", wave, registration = TRUE) %>%
  read_rds() %>%
  filter(status == "active") %>%
  mutate(age_group = age(date_of_birth, wave, grouped = TRUE)) %>%
  filter(gender %in% c("Male", "Female")) %>%
  group_by(age_group, gender) %>%
  summarise(n_reg = n(), .groups = "drop")

latest <-
  resp %>%
  full_join(reg, by = c("age_group", "gender")) %>%
  replace_na(list(n_resp = 0, n_reg = 0, n_opt_outs = 0)) %>%
  mutate(
    survey_wave = wave,
    date = start_date(wave)
  ) %>%
  select(survey_wave, date, everything())

```

### Responses

<br>

```{r latest-resp-table}

latest %>%
  select(-survey_wave, -date, -n_reg) %>%
  pivot_wider(names_from = age_group, values_from = n_resp) %>%
  adorn_totals("row") %>%
  adorn_totals("col") %>%
  rename(Gender = gender) %>%
  flextable(cwidth = 1.2) %>%
  theme_alafoli() %>%
  fontsize(size = 16, part = "all") %>%
  color(color = "black", part = "all") %>%
  padding(padding = 6, part = "all") %>%
  hline(i = 2, part = "body", border = fp_border(color = "black")) %>%
  vline(j = 7, part = "all", border = fp_border(color = "black")) %>%
  bold(j = 1, part = "body") %>%
  bold(part = "header") %>%
  bg(i = 3, j = 1:8, bg = "snow2", part = "body") %>%
  bg(j = 8, bg = "snow2", part = "all")

```

<br>

```{r latest-resp}

latest %>%
  ggplot(aes(x = age_group, y = n_resp, fill = gender)) +
  geom_col(position = dodge) +
  geom_text(aes(y = n_resp + 10, label = n_resp), 
            position = dodge,
            size = 5) +
  scale_fill_discrete_scs() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 300)) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        text = element_text(size = 16)) +
  labs(
    x = "Age Group",
    y = "Number of Responses",
    fill = "Gender"
  )

```

<br>

### Response Rate {.tabset}

#### Age Group

Bar labels indicate percentage response rate for each age group.

```{r latest-rate-age}

latest %>%
  group_by(age_group) %>%
  summarise(across(c(n_resp, n_reg), sum), .groups = "drop") %>%
  mutate(rate = paste0(round_half_up(n_resp / n_reg * 100, 0), "%"),
         total = n_reg,
         n_reg = n_reg - n_resp) %>%
  pivot_longer(cols = starts_with("n"), 
               names_to = "response", 
               values_to = "n") %>%
  mutate(response = ifelse(response == "n_resp", 
                           "Responded", "No Response")) %>%
  ggplot(aes(x = age_group, y = n)) +
  geom_col(aes(fill = as.character(response)), position = "stack") +
  geom_text(aes(y = total + 20, label = rate),
            size = 5) +
  scale_fill_discrete_scs() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 600)) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        text = element_text(size = 16)) +
  labs(
    x = "Age Group",
    y = "Number of Panel Members",
    fill = "Response Status"
  )

```

#### Gender

Bar labels indicate percentage response rate for each gender.

```{r latest-rate-gender}

latest %>%
  group_by(gender) %>%
  summarise(across(c(n_resp, n_reg), sum), .groups = "drop") %>%
  mutate(rate = paste0(round_half_up(n_resp / n_reg * 100, 0), "%"),
         total = n_reg,
         n_reg = n_reg - n_resp) %>%
  pivot_longer(cols = starts_with("n"), 
               names_to = "response", 
               values_to = "n") %>%
  mutate(response = ifelse(response == "n_resp", 
                           "Responded", "No Response")) %>%
  ggplot(aes(x = gender, y = n)) +
  geom_col(aes(fill = as.character(response)), position = "stack") +
  geom_text(aes(y = total + 40, label = rate),
            size = 5) +
  scale_fill_discrete_scs() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1500)) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        text = element_text(size = 16)) +
  labs(
    x = "Gender",
    y = "Number of Panel Members",
    fill = "Response Status"
  )

```

<br>

### Opt Outs

<br>

```{r latest-opt-outs-data, include = FALSE}

opt_outs <- opt_out_data(wave)

```

```{r latest-opt-outs-table}

opt_outs %>%
  select(-c(survey_wave, date)) %>%
  rename(`Age Group` = age_group, Gender = gender, `Opt Outs` = n) %>%
  adorn_totals(fill = "") %>%
  flextable(cwidth = 0.9) %>%
  width(j = 1:2, width = 1.8) %>%
  width(j = 3, width = 1.5) %>%
  theme_alafoli() %>%
  align(j = 1:2, align = "left", part = "all") %>%
  fontsize(size = 16, part = "all") %>%
  color(color = "black", part = "all") %>%
  padding(padding = 6, part = "all") %>%
  bold(part = "header") %>%
  hline(i = nrow(opt_outs), 
        part = "body", 
        border = fp_border(color = "black")) %>%
  bg(i = nrow(opt_outs) + 1, bg = "snow2", part = "body") %>%
  bold(i = nrow(opt_outs) + 1, part = "body")

```

<br>

## Trend

```{r data-trend, include = FALSE}

demog_data <- function(x) {

  get_resp <- function(y) {
    
    file <- suppressWarnings(
      scs_filepath("registration-data-anon", y, create_folder = FALSE)
    )
    
    if(!file.exists(file)) {
      stop("File doesn't exist: ", file)
    }
  
    read_rds(file) %>%
      mutate(age_group = age(date_of_birth,
                             as.numeric(str_extract(y, "\\d{1,2}")),
                             str_extract(y, "[AB]"),
                             grouped = TRUE)) %>%
      filter(gender %in% c("Male", "Female")) %>%
      group_by(age_group, gender) %>%
      summarise(n_resp = n(), .groups = "drop")
    
}
  
  if(!(x %in% 30:42) & !(x >= 44)) {
    stop("Wave must be greater or equal to 30, and not 43.")
  }

  data <- if(x %in% 30:42) {
    bind_rows(get_resp(paste0(x, "A")), get_resp(paste0(x, "B")))
  } else {
    get_resp(x)
  }

  data %>%
    group_by(age_group, gender) %>%
    summarise(across(everything(), sum), .groups = "drop") %>%
    mutate(
      survey_wave = x,
      date = start_date(as.numeric(x), "A")
    ) %>%
    select(survey_wave, date, everything())

}

safe_demog_data <- possibly(demog_data, NULL)

trend <- map_dfr(c(31:42, 44:wave), safe_demog_data)

```

### Responses {.tabset}

<br>

```{r trend-resp-table}

trend %>%
  filter(survey_wave >= 44) %>%
  group_by(survey_wave, date, age_group) %>%
  summarise(n_resp = sum(n_resp), .groups = "drop") %>%
  mutate(date = date_range(survey_wave, suffix = FALSE)) %>%
  rename(Wave = survey_wave, Date = date) %>%
  pivot_wider(names_from = age_group, values_from = n_resp) %>%
  adorn_totals("col") %>%
  flextable(cwidth = 0.9) %>%
  width(j = 2, width = 2.6) %>%
  theme_alafoli() %>%
  align(j = 1:2, align = "left", part = "all") %>%
  fontsize(size = 16, part = "all") %>%
  color(color = "black", part = "all") %>%
  padding(padding = 6, part = "all") %>%
  vline(j = 8, part = "all", border = fp_border(color = "black")) %>%
  bold(part = "header") %>%
  bg(j = 9, bg = "snow2", part = "all")

```

<br>

#### Overall

```{r trend-resp}

trend %>%
  group_by(date) %>%
  summarise(n = sum(n_resp), .groups = "drop") %>%
  ggplot(aes(x = date, y = n)) +
  geom_segment(aes(x = ymd("2021-11-25"), xend = ymd("2021-11-25"), 
                   y = 0, yend = 4000),
               colour = "#8e8e8e",
               linetype = "dashed") +
  geom_text(aes(x = ymd("2021-11-26"), y = 3800, label = "Panel\nRefresh"),
            hjust = 0, size = 5) +
  geom_rect(aes(xmin = ymd("2021-12-23"), xmax = ymd("2022-01-06"), 
                ymin = 0, ymax = 4000),
            fill = "#d4d4d4",
            alpha = 0.02) +
  geom_text(aes(x = ymd("2022-01-07"), y = 3900, label = "Festive Period"),
            hjust = 0, size = 5) +
  geom_line(colour = scs_colour_values["dark-blue"], size = 1) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 4200)) +
  scale_x_date(breaks = unique(trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Number of Responses"
  )

```

#### Age Group

```{r trend-resp-age}

trend %>%
  group_by(date, age_group) %>%
  summarise(n = sum(n_resp), .groups = "drop") %>%
  ggplot() +
  geom_segment(aes(x = ymd("2021-11-25"), xend = ymd("2021-11-25"), 
                   y = 0, yend = 1000),
               colour = "#8e8e8e",
               linetype = "dashed") +
  geom_text(aes(x = ymd("2021-11-26"), y = 960, label = "Panel\nRefresh"),
            hjust = 0, size = 5) +
  geom_rect(aes(xmin = ymd("2021-12-23"), xmax = ymd("2022-01-06"), 
                ymin = 0, ymax = 1000),
            fill = "#d4d4d4",
            alpha = 0.02) +
  geom_text(aes(x = ymd("2022-01-07"), y = 970, label = "Festive Period"),
            hjust = 0, size = 5) +
  geom_line(aes(x = date, y = n, colour = age_group), size = 1) +  
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1100)) +
  scale_x_date(breaks = unique(trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Number of Responses",
    colour = "Age Group"
  )

```

#### 18-29 Age Group

```{r trend-resp-18-29}

trend %>%
  filter(age_group == "18-29") %>%
  group_by(date) %>%
  summarise(n = sum(n_resp), .groups = "drop") %>%
  ggplot() +
  geom_segment(aes(x = ymd("2021-11-25"), xend = ymd("2021-11-25"), 
                   y = 0, yend = 220),
               colour = "#8e8e8e",
               linetype = "dashed") +
  geom_text(aes(x = ymd("2021-11-26"), y = 210, label = "Panel\nRefresh"),
            hjust = 0, size = 5) +
  geom_rect(aes(xmin = ymd("2021-12-23"), xmax = ymd("2022-01-06"), 
                ymin = 0, ymax = 220),
            fill = "#d4d4d4",
            alpha = 0.02) +
  geom_text(aes(x = ymd("2022-01-07"), y = 210, label = "Festive Period"),
            hjust = 0, size = 5) +
  geom_line(aes(x = date, y = n), 
            colour = scs_colour_values["dark-blue"], size = 1) +  
  geom_hline(aes(yintercept = 30), colour = "red", linetype = "dashed") +
  geom_text(aes(x = ymd("2022-04-20"), y = 25, label = "30 responses"), 
            colour = "red", hjust = 0, size = 5) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 230)) +
  scale_x_date(breaks = unique(trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Number of Responses"
  )

```

#### Gender

```{r trend-resp-gender}

trend %>%
  group_by(date, gender) %>%
  summarise(n = sum(n_resp), .groups = "drop") %>%
  ggplot() +
  geom_segment(aes(x = ymd("2021-11-25"), xend = ymd("2021-11-25"), 
                   y = 0, yend = 3000),
               colour = "#8e8e8e",
               linetype = "dashed") +
  geom_text(aes(x = ymd("2021-11-26"), y = 2800, label = "Panel\nRefresh"),
            hjust = 0, size = 5) +
  geom_rect(aes(xmin = ymd("2021-12-23"), xmax = ymd("2022-01-06"), 
                ymin = 0, ymax = 3000),
            fill = "#d4d4d4",
            alpha = 0.02) +
  geom_text(aes(x = ymd("2022-01-07"), y = 2900, label = "Festive Period"),
            hjust = 0, size = 5) +
  geom_line(aes(x = date, y = n, colour = gender), size = 1) +
  scale_colour_discrete_scs() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 3200)) +
  scale_x_date(breaks = unique(trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Number of Responses",
    colour = "Gender"
  )

```

### Response Rate {.tabset}

```{r trend-rate-data, include = FALSE}

# Full details of invited participants only available for previous 4 waves
# due to data retention policy.

reg <-
  list.files(here("survey-data", "registration-data"), full.names = TRUE) %>%
  map_dfr(
    ~ read_rds(.x) %>% 
      filter(status == "active") %>%
      mutate(wave = parse_number(basename(.x))) %>%
      mutate(age_group = age(date_of_birth, wave, grouped = TRUE)) %>%
      filter(gender %in% c("Male", "Female")) %>%
      group_by(wave, age_group, gender) %>%
      summarise(n_reg = n(), .groups = "drop") %>%
      mutate(date = start_date(wave)) %>%
      select(wave, date, everything())
  )

rate_trend <- 
  trend %>% 
  rename(wave = survey_wave) %>%
  filter(wave %in% reg$wave) %>%
  full_join(reg, by = c("wave", "date", "age_group", "gender"))

```

<br>

```{r trend-rate-table}

rate_trend %>%
  group_by(wave, date, age_group) %>%
  summarise(across(starts_with("n"), sum), .groups = "drop") %>%
  bind_rows(
    rate_trend %>%
      group_by(wave, date) %>%
      summarise(across(starts_with("n"), sum), .groups = "keep") %>%
      mutate(age_group = "Total") %>%
      ungroup()
  ) %>%
  arrange(wave) %>%
  mutate(rate = paste0(sprintf("%.1f", 
                               round_half_up(n_resp / n_reg * 100, 1)),
                       "%")) %>%
  select(-n_resp, -n_reg) %>%
  mutate(date = date_range(wave, suffix = FALSE)) %>%
  rename(Wave = wave, Date = date) %>%
  pivot_wider(names_from = age_group, values_from = rate) %>%
  flextable(cwidth = 0.9) %>%
  width(j = 2, width = 2.6) %>%
  theme_alafoli() %>%
  align(j = 1:2, align = "left", part = "all") %>%
  align(j = 3:9, align = "right", part = "all") %>%
  fontsize(size = 16, part = "all") %>%
  color(color = "black", part = "all") %>%
  padding(padding = 6, part = "all") %>%
  vline(j = 8, part = "all", border = fp_border(color = "black")) %>%
  bold(part = "header") %>%
  bg(j = 9, bg = "snow2", part = "all")

```

<br>

#### Overall

```{r trend-rate}

rate_trend %>%
  group_by(date) %>%
  summarise(across(starts_with("n"), sum), .groups = "drop") %>%
  mutate(rate = n_resp / n_reg * 100) %>%
  ggplot(aes(x = date, y = rate)) +
  geom_line(colour = scs_colour_values["dark-blue"], size = 1) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100),
                     labels = function(x) paste0(x, "%")) +
  scale_x_date(breaks = unique(rate_trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Response Rate"
  )

```

#### Age Group

```{r trend-rate-age}

rate_trend %>%
  group_by(date, age_group) %>%
  summarise(across(starts_with("n"), sum), .groups = "drop") %>%
  mutate(rate = n_resp / n_reg * 100) %>%
  ggplot(aes(x = date, y = rate, colour = age_group)) +
  geom_line(size = 1) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100),
                     labels = function(x) paste0(x, "%")) +
  scale_x_date(breaks = unique(rate_trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Response Rate",
    colour = "Age Group"
  )

```

### Opt Outs

```{r opt-outs-data, include = FALSE}

safe_opt_out_data <- possibly(opt_out_data, otherwise = NULL)

opt_outs <- map_dfr(c(31:42, 44:wave), safe_opt_out_data)

```

<br>

```{r opt-outs-table}

opt_outs %>%
  filter(survey_wave >= 44) %>%
  group_by(survey_wave, date) %>%
  summarise(n = sum(n), .groups = "drop") %>%
  mutate(date = date_range(survey_wave, suffix = FALSE)) %>%
  rename(Wave = survey_wave, Date = date, `Opt Outs` = n) %>%
  flextable(cwidth = 0.9) %>%
  width(j = 2, width = 2.6) %>%
  width(j = 3, width = 1.5) %>%
  theme_alafoli() %>%
  align(j = 1:2, align = "left", part = "all") %>%
  fontsize(size = 16, part = "all") %>%
  color(color = "black", part = "all") %>%
  padding(padding = 6, part = "all") %>%
  bold(part = "header")

```

<br>

```{r opt-outs}

opt_outs %>%
  group_by(date) %>%
  summarise(n = sum(n), .groups = "drop") %>%
  ggplot(aes(x = date, y = n)) +
  geom_segment(aes(x = ymd("2021-11-25"), xend = ymd("2021-11-25"),
                   y = 0, yend = 35),
               colour = "#8e8e8e",
               linetype = "dashed") +
  geom_text(aes(x = ymd("2021-11-26"), y = 32, label = "Panel\nRefresh"),
            hjust = 0, size = 5) +
  geom_rect(aes(xmin = ymd("2021-12-23"), xmax = ymd("2022-01-06"),
                ymin = 0, ymax = 35),
            fill = "#d4d4d4",
            alpha = 0.02) +
  geom_text(aes(x = ymd("2022-01-07"), y = 32, label = "Festive Period"),
            hjust = 0, size = 5) +
  geom_line(colour = scs_colour_values["dark-blue"], size = 1) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 35)) +
  scale_x_date(breaks = unique(trend$date), date_labels = "%d %b") +
  theme(text = element_text(size = 16)) +
  labs(
    x = "Date",
    y = "Number of Opt Outs"
  )

```